---
title: "Simulating a Continuous-Time, Stochastic, Double-Auction Market"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulating a Continuous-Time, Stochastic, Double-Auction Market}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(marketsimulatr)
```

To simulate an order book, the `create_order_book()` function is used. It takes as inputs the starting market price, the upper price limit and the tick size, as well as the arrival intensities of the various order types. The number of orders to produce is controlled using the `n_events` parameter.

The parameters can be specified manually, or by loading from the `config.yml` file.

```{r load_parameters}
# select config
Sys.setenv("R_CONFIG_ACTIVE" = "test")

# current price
start_price <- config::get("start_price")
# upper limit on price
price_limit <- config::get("price_limit")
# number of ticks from price within which limit orders are placed
price_band <- config::get("price_band")
# minimum price increment
tick_size <- config::get("tick_size")
# buy orders per minute
market_buy <- config::get("market_buy")
# sell orders per minute
market_sell <- config::get("market_sell")
# ratio of bids to market sells
bid_order_ratio <- config::get("bid_order_ratio")
# ratio of asks to market buys
ask_order_ratio <- config::get("ask_order_ratio")
# proportion on limit bids cancelled
cancel_bid <- config::get("cancel_bid")
# proportion of limit asks cancelled
cancel_ask <- config::get("cancel_ask")
# days to simulate
days <- config::get("days")
```

Next, the total number of events to simulate is calculated:

```{r calculate_num_events}
# this gives us a rough figure for orders per minute (of all types)
total_orders <- ceiling(
  market_buy +
    market_sell +
    (market_buy * ask_order_ratio) +
    (market_sell * bid_order_ratio) +
    (market_buy * ask_order_ratio * cancel_ask) +
    (market_sell * bid_order_ratio * cancel_bid)
)

# number of minutes to simulate
n_minutes <- days * 1440 # days * number of minutes in one day

# number of orders to generate (number of minutes * number of orders per minute)
n_events <- n_minutes * total_orders
```

Finally, the create_order_book function is called. This will return a price series of traded prices, as well as the full order book. The ask and bid limit order books are also returned.


```{r create_order_book}
simulation_results <- marketsimulatr::create_order_book(
  start_price,
  price_limit,
  price_band,
  tick_size,
  market_sell,
  market_buy,
  bid_order_ratio,
  ask_order_ratio,
  cancel_bid,
  cancel_ask,
  n_events
)
```
